RIScatter: Unifying Backscatter Communication and Reconfigurable Intelligent Surface

### 2023-08-15 main.tex
:triangular_flag_on_post: JSAC 230815

@@

% algpseudocode
\makeatletter
\renewcommand{\fnum@algorithm}{\fname@algorithm{} \thealgorithm:}
\makeatother

@@

\begin{section}{Introduction}
	\label{sc:introduction}
	\IEEEPARstart{F}{uture} wireless network is envisioned to provide high throughput, uniform coverage, pervasive connectivity, heterogeneous control, and cognitive intelligence for trillions of low-power devices.
	\gls{bc} separates a transmitter into a \gls{rf} carrier emitter with power-hungry elements (e.g., synthesizer and amplifier) and an information-bearing node with power-efficient components (e.g., harvester and modulator) \cite{Boyer2014}.
	The receiver (reader) can be either co-located or separated with the carrier emitter, known as \gls{mbc} and \gls{bbc} in Fig.~\subref*{fg:mbc} and \subref*{fg:bbc}, respectively.
	Relevant applications such as \gls{rfid} \cite{Dobkin2012,Landt2005} and passive sensor network \cite{Vannucci2008,Assimonis2016} have been extensively researched, standardized, and commercialized to embrace the \gls{ioe}.
	However, conventional backscatter nodes only respond when externally inquired by a nearby reader.
	\gls{ambc} in Fig.~\subref*{fg:ambc} was proposed a decade ago where battery-free nodes recycle ambient signals (e.g., radio, television and Wi-Fi) to harvest energy and establish connections \cite{Liu2013b}.
	It does not require dedicated power source, carrier emitter, or frequency spectrum, but the backscatter decoding is subject to the strong interference from the primary (legacy) link.
	To tackle this, cooperative \gls{ambc} \cite{Yang2018} employs a co-located receiver to decode both coexisting links, and the concept was further refined as \gls{sr} in Fig.~\subref*{fg:sr} \cite{Liang2020}.
	Specifically, the active transmitter generates \gls{rf} wave carrying primary information, the passive node creates a rich-scattering environment and rides its own information, and the co-located receiver cooperatively decodes both links.
	In those \gls{bc} applications, the scatter node is considered as an \emph{information source} and the reflection pattern depends exclusively on the information symbol.
	On the other hand, \gls{ris} in Fig.~\subref*{fg:ris} is a smart signal reflector with numerous passive elements of adjustable phase shifts.
	It customizes the wireless environment for signal enhancement, interference suppression, scattering enrichment, and/or non-line-of-sight bypassing \cite{Wu2021b}.
	Each \gls{ris} element is considered as a \emph{channel shaper} and the reflection pattern depends exclusively on the \gls{csi}.

@@

	On the other hand, static \gls{ris} that employs fixed reflection pattern per channel block has been extensively studied in wireless communication, sensing, and power literature \cite{Wu2018,Zhang2019a,Lin2022,Liu2022,Feng2022,Zhao2022}.
	Dynamic \gls{ris} performs time sharing between different phase shifts and introduces artificial channel diversity within each channel block.
	The idea was first proposed to fine-tune the \gls{ofdm} resource blocks \cite{Yang2020}, then extended to the downlink power and uplink information phases of \gls{wpcn} \cite{Wu2021,Wu2021d,Hua2022a}.
	However, dynamic \gls{ris} carries no information because the reflection state at a specific time is known to the receiver.
	\gls{ris} can also be used as an information source, and prototypes have been developed for \gls{psk} \cite{Tang2019a} and \gls{qam} \cite{Dai2020a}.
	From an information-theoretic perspective, the authors of \cite{Karasik2020} reported that joint transmitter-\gls{ris} encoding achieves the capacity of \gls{ris}-aided finite-input channel, and using \gls{ris} as a naive passive beamformer to maximize the receive \gls{snr} is generally suboptimal.
	This inspired \cite{Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hu2021b,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} to combine passive beamforming and backscatter modulation in the overall \gls{ris} design.
	In particular, \emph{symbol level precoding} maps the information symbols to the optimized \gls{ris} coefficient sets \cite{Liu2019d,Bereyhi2020}, \emph{overlay modulation} superposes the information symbols over a common auxiliary matrix \cite{Xu2020b,Zhang2021d,Hu2021b,Hua2022}, \emph{spatial modulation} switches between the reflection coefficient sets that maximize \gls{snr} at different receive antennas \cite{Basar2020,Ma2020a,Yuan2021}, and \emph{index modulation} employs dedicated reflection elements (resp. information elements) for passive beamforming (resp. backscatter modulation) \cite{Hu2021a}.
	Those \gls{ris}-based backscatter modulation schemes incur advanced hardware architecture and high optimization complexity.
	In contrast, \cite{Vardakis2023} exploits commodity \gls{rfid} tags, powered and controlled by a software-defined radio reader at a different frequency, to perform passive beamforming (but no backscatter modulation) towards a legacy user. \label{cm:2.5}
	% The authors demonstrated the feasibility of using batteryless devices to adaptively enhance the \gls{snr} of a primary link, but did not consider backscatter modulation at.
	Most relevant literature considers either Gaussian codebook \cite{Guo2019b,Ding2020,Long2020a,Zhou2019a,Wu2021a,Xu2021a,Yang2021a,Hu2021b} that is impractical for low-power nodes, or finite equiprobable inputs \cite{Yang2018,Liang2020,Han2021,Zhang2022,Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} that does not fully exploit the \gls{csi} and properties of active-passive coexisting networks.
	Those problems are addressed in this paper and the contributions are summarized below.

@@

	\emph{Third,} we address the second challenge in a single-user multi-node \gls{miso} scenario.
	We characterize the achievable primary-(total-)backscatter rate region by optimizing the input distribution at RIScatter nodes, the active beamforming at the \gls{ap}, and the energy decision regions at the user under different \gls{qos}.
	A \gls{bcd} algorithm is proposed where the \gls{kkt} input distribution is numerically evaluated by limit of sequences, the active beamforming is optimized by \gls{pga}, and the decision regions are refined by state-of-the-art sequential quantizer designs for \gls{dmtc}.
	Uniquely, our optimization problem takes into account the \gls{csi}, \gls{qos}, and backscatter constellation, and the resulting input distribution is applicable to any detection scheme.
	This is also the first paper to reveal the importance of backscatter input distribution and decision region designs in active-passive coexisting networks.

@@

\begin{section}{RIScatter}
	\label{sc:riscatter}
	\begin{subsection}{Principles}
		\label{sc:principles}
		\gls{rf} wave scattering or reflecting are often manipulated by passive antennas or programmable metamaterial \cite{Liang2022}.
		The former receives the impinging signals and reradiates some back to the space, while the latter reflects at the space-cell boundary and mainly applies a phase shift.
		In the scattered signal, the structural mode component depends on the scatterer geometry and material.
		Its impact is usually modelled as part of environment multipath \cite{Thomas2012,Liang2020}, or simply a baseband \gls{dc} offset when the impinging signal is a \gls{cw} \cite{Boyer2014}.
		On the other hand, the antenna mode component depends on the impedance mismatch and is widely exploited in scattering applications.
		For an antenna (resp. metamaterial) scatterer with $M$ reflection states, the reflection coefficient at state $m \in \mathcal{M} \triangleq \{1,\ldots,M\}$ is \cite{VanHuynh2017,Liang2022}
		\begin{equation}
			\Gamma_m = \frac{Z_m - Z^*}{Z_m + Z},
			\label{eq:reflection_coefficient}
		\end{equation}
		where $Z_m$ is the antenna load (resp. metamaterial cell) impedance at state $m$, and $Z$ is the antenna input (resp. medium characteristic) impedance.
		Specifically,
		\begin{itemize}
			\item \gls{bc}: The scatterer is an information source with random reflection pattern over time.
			The reflection coefficient is used merely an information codeword \cite{Thomas2012a}
			\begin{equation}
				\Gamma_m = \alpha_m \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
				\label{eq:backscatter_modulation}
			\end{equation}
			where $\alpha_m \in \mathbb{I}$ is the amplitude scattering ratio at state $m$, and $c_m$ is the corresponding constellation point.
			\item \gls{ris}: The scatterer is a channel shaper with deterministic reflection pattern over time.
			The reflection coefficient is used merely a passive beamforming codeword \cite{Wu2018}
			\begin{equation}
				\Gamma_m = \alpha_m \exp(j \theta_m),
				\label{eq:passive_beamforming}
			\end{equation}
			where $\theta_m$ is the phase shift at state $m$.\footnote{Most papers assume $\alpha_m = \alpha$, with $\alpha \ll 1$ for \gls{bc} and $\alpha=1$ for \gls{ris}.\label{fn:scattering_ratio}}
		\end{itemize}

@@

	\begin{subsection}{System Model}
		\label{sc:system_model}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{0.7\columnwidth}
			\footnotesize{
				\import{assets/illustration/}{riscatter_network.pdf_tex}
			}
			\caption{A single-user multi-node RIScatter network.}
			\label{fg:riscatter_network}
		\end{figure}
		As shown in Fig.~\ref{fg:riscatter_network}, we consider a RIScatter network where a $Q$-antenna \gls{ap} serves a single-antenna user and $K$ nearby dispersed or co-located RIScatter nodes.
		Without loss of generality, we assume all nodes have $M$ available reflection states.
		In the primary point-to-point system, the \gls{ap} transmits information to the user over a multipath channel\footnote{It is assumed the primary symbol duration is much longer than multipath delay spread (i.e., no inter-symbol interference).} enhanced by RIScatter nodes.
		In the backscatter multiple access system, the \gls{ap} acts as a carrier emitter, the RIScatter nodes modulate over the scattered signal, and the user jointly decodes all nodes.\footnote{It is assumed the signal going through two or more RIScatter nodes is too weak to be received by the user.}
		For simplicity, we consider a quasi-static block fading model and focus on a specific channel block where the \gls{csi} remains constant.
		Denote the \gls{ap}-user direct channel as $\boldsymbol{h}_{\text{D}}^\mathsf{H} \in \mathbb{C}^{1 \times Q}$, the \gls{ap}-node $k \in \mathcal{K} \triangleq \{1,\ldots,K\}$ forward channel as $\boldsymbol{h}_{\text{F},k}^\mathsf{H} \in \mathbb{C}^{1 \times Q}$, the node $k$-user backward channel as $h_{\text{B},k}$, and the cascaded \gls{ap}-node $k$-user channel as $\boldsymbol{h}_{\text{C},k}^\mathsf{H} \triangleq h_{\text{B},k} \boldsymbol{h}_{\text{F},k}^\mathsf{H} \in \mathbb{C}^{1 \times Q}$.
		We assume the direct and cascaded \gls{csi} are available at the \gls{ap} and user\footnote{The cascaded \gls{csi} can be estimated by sequential \cite{Bharadia2015,Yang2015b,Guo2019g} or parallel \cite{Jin2021a} approaches for dispersed nodes, or group-based \cite{Zheng2019} or hierarchical \cite{You2019} approaches for co-located nodes.}.

		Let $\alpha_k = \in \mathbb{I}$ be the amplitude scattering ratio of node $k$, $x_k \in \mathcal{X} \triangleq \{c_1,\ldots,c_M\}$ be the (coded) backscatter symbol of node $k$, and $x_{\mathcal{K}} \triangleq (x_1,\ldots,x_K)$ be the backscatter symbol tuple of all nodes.
		Due to double modulation, the composite channel is a function of backscatter symbol tuple\footnote{\eqref{eq:equivalent_channel_bc} and \eqref{eq:equivalent_channel_ris} are often used in \gls{bc} and \gls{ris} literature, respectively.}
		\begin{subequations}
			\label{eq:equivalent_channel}
			\begin{align}
				\boldsymbol{h}^\mathsf{H}(x_{\mathcal{K}})
				 & \triangleq \boldsymbol{h}_{\text{D}}^\mathsf{H} + \sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k \label{eq:equivalent_channel_bc}                    \\
				 & = \boldsymbol{h}_{\text{D}}^\mathsf{H} + \boldsymbol{x}^\mathsf{H} \mathrm{diag}(\boldsymbol{\alpha}) \boldsymbol{H}_{\text{C}}, \label{eq:equivalent_channel_ris}
			\end{align}
		\end{subequations}
		where $\boldsymbol{\alpha} \triangleq [\alpha_1,\ldots,\alpha_K]^\mathsf{T} \in \mathbb{I}^{K}$, $\boldsymbol{x} \triangleq [x_1,\ldots,x_K]^\mathsf{H} \in \mathcal{X}^{K}$, and $\boldsymbol{H}_{\text{C}} \triangleq [\boldsymbol{h}_{\text{C},1},\ldots,\boldsymbol{h}_{\text{C},K}]^\mathsf{H} \in \mathbb{C}^{K \times Q}$.
		Without loss of generality, we assume the spreading factor $N$ is a positive integer.
		Within one backscatter block, the signal received by the user at primary block $n \in \mathcal{N} \triangleq \{1,\ldots,N\}$ is
		\begin{equation}
			y[n] = \boldsymbol{h}^\mathsf{H}(x_{\mathcal{K}}) \boldsymbol{w} s[n] + v[n],
			\label{eq:receive_signal}
		\end{equation}
		where $\boldsymbol{w} \in \mathbb{C}^{Q}$ is the active beamformer satisfying $\lVert \boldsymbol{w} \rVert^2 \le P$, $P$ is the maximum average transmit power, $s \sim \mathcal{CN}(0,1)$ is the primary symbol, and $v \sim \mathcal{CN}(0,\sigma_v^2)$ is the \gls{awgn} with variance $\sigma_v^2$.

@@

		Let $L = M^K$ be the number of decision regions.
		Sort $\{\sigma_{m_{\mathcal{K}}}^2\}$ in ascending order and denote the result sequence as $\sigma_1^2,\ldots,\sigma_L^2$.
		With sequential thresholding, the decision region of backscatter symbol tuple $l \in \mathcal{L} = \{1,\ldots,L\}$ is\footnote{$m_{\mathcal{K}}$ and $l$ are one-to-one mapped. Both notations are used interchangeably in the following context.}
		\begin{equation}
			\mathcal{R}_{l} \triangleq [t_{l-1},t_l), \quad 0 \le t_{l-1} \le t_l,
		\end{equation}
		where $t_l$ is the decision threshold between hypotheses $x_{l}$ and $x_{l+1}$.
		An example is shown in Fig.~\ref{fg:energy_distribution}.

@@

\begin{section}{Rate-Region Characterization}
	With a slight abuse of notation, we define the weighed sum mutual information and information function as
	\begin{align}
		I(x_{\mathcal{K}})
		 & \triangleq \rho I_{\text{P}}(s;y|x_{\mathcal{K}}) + (1 - \rho) I_{\text{B}}(x_{\mathcal{K}};\hat{x}_{\mathcal{K}}),\label{eq:weighted_sum_mutual_information}           \\
		I(x_{m_\mathcal{K}})
		 & \triangleq \rho I_{\text{P}}(s;y|x_{m_{\mathcal{K}}}) + (1 - \rho) I_{\text{B}}(x_{m_{\mathcal{K}}};\hat{x}_{\mathcal{K}}),\label{eq:weighted_sum_information_function}
	\end{align}
	where $\rho \in \mathbb{I}$ is the \gls{qos}.
	To obtain the achievable primary-(total-)backscatter rate region, we consider the weighted sum mutual information maximization problem with independent encoding at all nodes\footnote{Joint encoding over multiple nodes can be viewed as the special case with an augmented source.\label{fn:joint_encoding}}
	\begin{maxi!}
		{\scriptstyle{\{\boldsymbol{p}_k\}_{k \in \mathcal{K}},\boldsymbol{w},\boldsymbol{t}}}{I(x_{\mathcal{K}})}{\label{op:weighted_sum_rate}}{\label{ob:weighted_sum_rate}}
		\addConstraint{\boldsymbol{1}^\mathsf{T} \boldsymbol{p}_k}{=1,}{\quad \forall k}{\label{co:sum_probability}}
		\addConstraint{\boldsymbol{p}_k}{\ge \boldsymbol{0},}{\quad \forall k}{\label{co:nonnegative_probability}}
		\addConstraint{\lVert \boldsymbol{w} \rVert^2}{\le P}{\label{co:transmit_power}}
		\addConstraint{t_{l-1}}{\le t_l,}{\quad \forall l}{\label{co:sequential_threshold}}
		\addConstraint{\boldsymbol{t}}{\ge \boldsymbol{0},}{\label{co:nonnegative_threshold}}
	\end{maxi!}
	where $\boldsymbol{p}_k = [p(x_{1_k}), \ldots, p(x_{M_k})]^\mathsf{T} \in \mathbb{I}^M$ is the input distribution of node $k$.
	Problem \eqref{op:weighted_sum_rate} generalizes \gls{bc} by allowing \gls{csi}- and \gls{qos}-adaptive input distribution and decision region design.
	On the other hand, it also relaxes the feasible domain of discrete \gls{ris} phase shift selection problem from vertices of $M$-dimensional probability simplex to the simplex itself.
	The original problem is highly non-convex and we propose a \gls{bcd} algorithm that iteratively updates $\{\boldsymbol{p}_k\}_{k \in \mathcal{K}}$, $\boldsymbol{w}$ and $\boldsymbol{t}$.

>>>>

% page limit hacks
% \usepackage{setspace}
% ! \usepackage[top=1cm, bottom=1cm, left=1cm, right=1cm]{geometry}
% \abovedisplayskip=1mm
% \belowdisplayskip=1mm
% \abovedisplayshortskip=1mm
% \belowdisplayshortskip=1mm
% \setlength{\jot}{0.1mm}
% \setlength{\floatsep}{1mm}
% \setlength{\textfloatsep}{1mm}
% \setlength{\intextsep}{1mm}
% \setlength{\skip\footins}{2mm}

@@

% algpseudocode
\makeatletter
\renewcommand{\fnum@algorithm}{\fname@algorithm{} \thealgorithm:}
\newcommand\setalgorithmcaptionfont[1]{%
	\let\my@floatc@ruled\floatc@ruled          % save \floatc@ruled
	\def\floatc@ruled{%
		\global\let\floatc@ruled\my@floatc@ruled % restore \floatc@ruled
		#1\floatc@ruled}}
\makeatother

@@

\begin{section}{Introduction}
	\label{sc:introduction}
	\IEEEPARstart{F}{uture} wireless network is envisioned to provide high throughput, uniform coverage, pervasive connectivity, heterogeneous control, and cognitive intelligence for trillions of low-power devices.
	\gls{bc} separates a transmitter into a \gls{rf} carrier emitter with power-hungry elements (e.g., synthesizer and amplifier) and an information-bearing node with power-efficient components (e.g., harvester and modulator) \cite{Boyer2014}.
	The receiver (reader) can be either co-located or separated with the carrier emitter, known as \gls{mbc} and \gls{bbc} in Fig.~\subref*{fg:mbc} and \subref*{fg:bbc}, respectively.
	Relevant applications such as \gls{rfid} \cite{Dobkin2012,Landt2005} and passive sensor network \cite{Vannucci2008,Assimonis2016} have been extensively researched, standardized, and commercialized to embrace the \gls{ioe}.
	However, conventional backscatter nodes only respond when externally inquired by a nearby reader.
	\gls{ambc} in Fig.~\subref*{fg:ambc} was proposed a decade ago where battery-free nodes recycle ambient signals (e.g., radio, television and Wi-Fi) to harvest energy and establish connections \cite{Liu2013b}.
	It does not require dedicated power source, carrier emitter, or frequency spectrum, but the backscatter decoding is subject to the strong interference from the primary (legacy) link.
	To tackle this, cooperative \gls{ambc} \cite{Yang2018} employs a co-located receiver to decode both coexisting links, and the concept was further refined as \gls{sr} in Fig.~\subref*{fg:sr} \cite{Liang2020}.
	Specifically, the active transmitter generates \gls{rf} wave carrying primary information, the passive node creates a rich-scattering environment and superimposes its own information, and the co-located receiver cooperatively decodes both links.
	In those \gls{bc} applications, the scatter node is considered as an \emph{information source} and the reflection pattern depends exclusively on the information symbol.
	On the other hand, \gls{ris} in Fig.~\subref*{fg:ris} is a smart signal reflector with numerous passive elements of adjustable phase shifts.
	It customizes the wireless environment for signal enhancement, interference suppression, scattering enrichment, and/or non-line-of-sight bypassing \cite{Wu2021b}.
	Each \gls{ris} element is considered as a \emph{channel shaper} and the reflection pattern depends exclusively on the \gls{csi}.

@@

	On the other hand, static \gls{ris} that employs fixed reflection pattern per channel block has been extensively studied in wireless communication, sensing, and power literature \cite{Wu2018,Zhang2019a,Lin2022,Liu2022,Feng2022,Zhao2022}.
	Dynamic \gls{ris} performs time sharing between different phase shifts and introduces artificial channel diversity within each channel block.
	The idea was first proposed to fine-tune the \gls{ofdm} resource blocks \cite{Yang2020}, then extended to the downlink power and uplink information phases of \gls{wpcn} \cite{Wu2021,Wu2021d,Hua2022a}.
	However, dynamic \gls{ris} carries no information because the reflection state at a specific time is known to the receiver.
	\gls{ris} can also be used as an information source, and prototypes have been developed for \gls{psk} \cite{Tang2019a} and \gls{qam} \cite{Dai2020a}.
	From an information-theoretic perspective, the authors of \cite{Karasik2020} reported that joint transmitter-\gls{ris} encoding achieves the capacity of \gls{ris}-aided finite-input channel, and using \gls{ris} as a naive passive beamformer to maximize the receive \gls{snr} is generally rate-suboptimal.
	This inspired \cite{Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hu2021b,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} to combine passive beamforming and backscatter modulation in the overall \gls{ris} design.
	In particular, \emph{symbol level precoding} maps the information symbols to the optimized \gls{ris} coefficient sets \cite{Liu2019d,Bereyhi2020}, \emph{overlay modulation} superposes the information symbols over a common auxiliary matrix \cite{Xu2020b,Zhang2021d,Hu2021b,Hua2022}, \emph{spatial modulation} switches between the reflection coefficient sets that maximize \gls{snr} at different receive antennas \cite{Basar2020,Ma2020a,Yuan2021}, and \emph{index modulation} employs dedicated reflection elements (resp. information elements) for passive beamforming (resp. backscatter modulation) \cite{Hu2021a}.
	Those \gls{ris}-based backscatter modulation schemes incur advanced hardware architecture and high optimization complexity.
	In contrast, \cite{Vardakis2023} exploits commodity \gls{rfid} tags, powered and controlled by a software-defined radio reader at a different frequency, to perform passive beamforming (but no backscatter modulation) towards a legacy user. \label{cm:2.5}
	% The authors demonstrated the feasibility of using batteryless devices to adaptively enhance the \gls{snr} of a primary link, but did not consider backscatter modulation at.
	Most relevant literature considers either Gaussian codebook \cite{Guo2019b,Ding2020,Long2020a,Zhou2019a,Wu2021a,Xu2021a,Yang2021a,Hu2021b} that is impractical for low-power nodes, or finite equiprobable inputs \cite{Yang2018,Liang2020,Han2021,Zhang2022,Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} that does not fully exploit the \gls{csi} and properties of active-passive coexisting networks.
	Those problems are addressed in this paper and the contributions are summarized below.

@@

	\emph{Third,} we address the second challenge in a single-user multi-node \gls{miso} scenario.
	We characterize the achievable primary-(total-)backscatter rate region by optimizing the input distribution at RIScatter nodes, the active beamforming at the \gls{ap}, and the energy decision regions at the user under different \gls{qos}.
	A \gls{bcd} algorithm is proposed where the \gls{kkt} input distribution is numerically evaluated by the converging point of a sequence, the active beamforming is optimized by \gls{pga}, and the decision regions are refined by state-of-the-art sequential quantizer designs for \gls{dmtc}.
	Uniquely, our optimization problem takes into account the \gls{csi}, \gls{qos}, and backscatter constellation, and the resulting input distribution is applicable to other detection schemes.
	This is also the first paper to reveal the importance of backscatter input distribution and decision region designs in active-passive coexisting networks.

@@

\begin{section}{RIScatter}
	\label{sc:riscatter}
	\begin{subsection}{Principles}
		\label{sc:principles}
		\gls{rf} wave scattering or reflecting are often manipulated by passive antennas or programmable metamaterial \cite{Liang2022}.
		The former receives the impinging signals and reradiates some back to the space, while the latter reflects at the space-cell boundary and mainly applies a phase shift.
		In the scattered signal, the structural mode component depends on the scatterer geometry and material.
		Its impact is usually modelled as part of environment multipath \cite{Thomas2012,Liang2020}, or simply a baseband \gls{dc} offset when the impinging signal is a \gls{cw} \cite{Boyer2014}.
		On the other hand, the antenna mode component depends on the impedance mismatch and is widely exploited in scattering applications.
		For an antenna (resp. metamaterial) scatterer with $M$ reflection states, the reflection coefficient at state $m \in \mathcal{M} \triangleq \{1,\ldots,M\}$ is \cite{VanHuynh2017,Liang2022}
		\begin{equation}
			\Gamma_m = \frac{Z_m - Z^*}{Z_m + Z},
			\label{eq:reflection_coefficient}
		\end{equation}
		where $Z_m$ is the antenna load (resp. metamaterial cell) impedance at state $m$, and $Z$ is the antenna input (resp. medium characteristic) impedance.
		Specifically,
		\begin{itemize}
			\item \gls{bc}: The scatterer is an information source with random reflection pattern over time.
			The reflection coefficient is used merely as an information codeword \cite{Thomas2012a}
			\begin{equation}
				\Gamma_m = \alpha_m \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
				\label{eq:backscatter_modulation}
			\end{equation}
			where $\alpha_m \in \mathbb{I}$ is the amplitude scattering ratio at state $m$, and $c_m$ is the corresponding constellation point.
			\item \gls{ris}: The scatterer is a channel shaper with deterministic reflection pattern over time.
			The reflection coefficient is used merely as a passive beamforming codeword \cite{Wu2018}
			\begin{equation}
				\Gamma_m = \alpha_m \exp(j \theta_m),
				\label{eq:passive_beamforming}
			\end{equation}
			where $\theta_m$ is the phase shift at state $m$.\footnote{Most papers assume $\alpha_m = \alpha$, with $\alpha \ll 1$ for \gls{bc} and $\alpha=1$ for \gls{ris}.\label{fn:scattering_ratio}}
		\end{itemize}

@@

	\begin{subsection}{System Model}
		\label{sc:system_model}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{0.7\columnwidth}
			\footnotesize{
				\import{assets/illustration/}{riscatter_network.pdf_tex}
			}
			\caption{A single-user multi-node RIScatter network.}
			\label{fg:riscatter_network}
		\end{figure}
		As shown in Fig.~\ref{fg:riscatter_network}, we consider a RIScatter network where a $Q$-antenna \gls{ap} serves a single-antenna user and $K$ nearby dispersed or co-located RIScatter nodes.
		Without loss of generality, we assume all nodes have $M$ available reflection states.
		In the primary point-to-point system, the \gls{ap} transmits information to the user over a multipath channel\footnote{It is assumed the primary symbol duration is much longer than multipath delay spread (i.e., no inter-symbol interference).} enhanced by RIScatter nodes.
		In the backscatter multiple access system, the \gls{ap} acts as a carrier emitter, the RIScatter nodes modulate over the scattered signal, and the user jointly decodes all nodes.\footnote{It is assumed the signal going through two or more RIScatter nodes is too weak to be received by the user.}
		For simplicity, we consider a quasi-static block fading model and focus on a specific channel block where the \gls{csi} remains constant.
		Denote the \gls{ap}-user direct channel as $\boldsymbol{h}_{\text{D}}^\mathsf{H} \in \mathbb{C}^{1 \times Q}$, the \gls{ap}-node $k \in \mathcal{K} \triangleq \{1,\ldots,K\}$ forward channel as $\boldsymbol{h}_{\text{F},k}^\mathsf{H} \in \mathbb{C}^{1 \times Q}$, the node $k$-user backward channel as $h_{\text{B},k}$, and the cascaded \gls{ap}-node $k$-user channel as $\boldsymbol{h}_{\text{C},k}^\mathsf{H} \triangleq h_{\text{B},k} \boldsymbol{h}_{\text{F},k}^\mathsf{H} \in \mathbb{C}^{1 \times Q}$.
		We assume the direct and cascaded \gls{csi} are available at the \gls{ap} and user\footnote{The cascaded \gls{csi} can be estimated by sequential \cite{Bharadia2015,Yang2015b,Guo2019g} or parallel \cite{Jin2021a} approaches for dispersed nodes, or group-based \cite{Zheng2019} or hierarchical \cite{You2019} approaches for co-located nodes. The impact of channel estimation error will be investigated in Section~\ref{sc:simulation_results}.}.

		Let $\alpha_k \in \mathbb{I}$ be the amplitude scattering ratio of node $k$, $x_k \in \mathcal{X} \triangleq \{c_1,\ldots,c_M\}$ be the (coded) backscatter symbol of node $k$, and $x_{\mathcal{K}} \triangleq (x_1,\ldots,x_K)$ be the backscatter symbol tuple of all nodes.
		Due to double modulation, the composite channel is a function of backscatter symbol tuple\footnote{\eqref{eq:equivalent_channel_bc} and \eqref{eq:equivalent_channel_ris} are often used in \gls{bc} and \gls{ris} literature, respectively.}
		\begin{subequations}
			\label{eq:equivalent_channel}
			\begin{align}
				\boldsymbol{h}^\mathsf{H}(x_{\mathcal{K}})
				 & \triangleq \boldsymbol{h}_{\text{D}}^\mathsf{H} + \sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k \label{eq:equivalent_channel_bc}                    \\
				 & = \boldsymbol{h}_{\text{D}}^\mathsf{H} + \boldsymbol{x}^\mathsf{H} \mathrm{diag}(\boldsymbol{\alpha}) \boldsymbol{H}_{\text{C}}, \label{eq:equivalent_channel_ris}
			\end{align}
		\end{subequations}
		where $\boldsymbol{\alpha} \triangleq [\alpha_1,\ldots,\alpha_K]^\mathsf{T} \in \mathbb{I}^{K}$, $\boldsymbol{x} \triangleq [x_1,\ldots,x_K]^\mathsf{H} \in \mathcal{X}^{K}$, and $\boldsymbol{H}_{\text{C}} \triangleq [\boldsymbol{h}_{\text{C},1},\ldots,\boldsymbol{h}_{\text{C},K}]^\mathsf{H} \in \mathbb{C}^{K \times Q}$.
		Without loss of generality, we assume the spreading factor $N$ is a positive integer.
		Within one backscatter block, the signal received by the user at primary block $n \in \mathcal{N} \triangleq \{1,\ldots,N\}$ is
		\begin{equation}
			y[n] = \boldsymbol{h}^\mathsf{H}(x_{\mathcal{K}}) \boldsymbol{w} s[n] + v[n],
			\label{eq:receive_signal}
		\end{equation}
		where $\boldsymbol{w} \in \mathbb{C}^{Q}$ is the active beamformer satisfying $\lVert \boldsymbol{w} \rVert^2 \le P$, $P$ is the maximum average transmit power, $s \sim \mathcal{CN}(0,1)$ is the primary symbol, and $v \sim \mathcal{CN}(0,\sigma_v^2)$ is the \gls{awgn} with variance $\sigma_v^2$.

@@

		Let $L = M^K$ be the number of decision regions.
		Sort $\{\sigma_{m_{\mathcal{K}}}^2\}$ in ascending order and denote the result sequence as $\sigma_1^2,\ldots,\sigma_L^2$.
		With sequential thresholding, the decision region of backscatter symbol tuple $l \in \mathcal{L} \triangleq \{1,\ldots,L\}$ is\footnote{$m_{\mathcal{K}}$ and $l$ are one-to-one mapped. Both notations are used interchangeably in the following context.}
		\begin{equation}
			\mathcal{R}_{l} \triangleq [t_{l-1},t_l), \quad 0 \le t_{l-1} \le t_l,
		\end{equation}
		where $t_l$ is the decision threshold between hypotheses $x_{l}$ and $x_{l+1}$.
		An example is shown in Fig.~\ref{fg:energy_distribution}.

@@

\begin{section}{Rate-Region Characterization}
	With a slight abuse of notation, we define the weighed sum mutual information and information function as
	\begin{align}
		I(x_{\mathcal{K}})
		 & \triangleq \rho I_{\text{P}}(s;y|x_{\mathcal{K}}) + (1 - \rho) I_{\text{B}}(x_{\mathcal{K}};\hat{x}_{\mathcal{K}}),\label{eq:weighted_sum_mutual_information}           \\
		I(x_{m_\mathcal{K}})
		 & \triangleq \rho I_{\text{P}}(s;y|x_{m_{\mathcal{K}}}) + (1 - \rho) I_{\text{B}}(x_{m_{\mathcal{K}}};\hat{x}_{\mathcal{K}}),\label{eq:weighted_sum_information_function}
	\end{align}
	where $\rho \in \mathbb{I}$ is the \gls{qos}.
	To obtain the achievable primary-(total-)backscatter rate region, we consider the weighted sum mutual information maximization problem with independent encoding at all nodes\footnote{Joint encoding over multiple nodes can be viewed as its special case with an augmented backscatter source.\label{fn:joint_encoding}}
	\begin{maxi!}
		{\scriptstyle{\{\boldsymbol{p}_k\}_{k \in \mathcal{K}},\boldsymbol{w},\boldsymbol{t}}}{I(x_{\mathcal{K}})}{\label{op:weighted_sum_rate}}{\label{ob:weighted_sum_rate}}
		\addConstraint{\boldsymbol{1}^\mathsf{T} \boldsymbol{p}_k}{=1,}{\quad \forall k}{\label{co:sum_probability}}
		\addConstraint{\boldsymbol{p}_k}{\ge \boldsymbol{0},}{\quad \forall k}{\label{co:nonnegative_probability}}
		\addConstraint{\lVert \boldsymbol{w} \rVert^2}{\le P}{\label{co:transmit_power}}
		\addConstraint{t_{l-1}}{\le t_l,}{\quad \forall l}{\label{co:sequential_threshold}}
		\addConstraint{\boldsymbol{t}}{\ge \boldsymbol{0},}{\label{co:nonnegative_threshold}}
	\end{maxi!}
	where $\boldsymbol{p}_k = [p(x_{1_k}), \ldots, p(x_{M_k})]^\mathsf{T} \in \mathbb{I}^M$ is the input distribution of node $k$.
	Problem \eqref{op:weighted_sum_rate} generalizes \gls{bc} by allowing \gls{csi}- and \gls{qos}-adaptive input distribution and decision region design.
	On the other hand, it also relaxes the feasible domain of discrete \gls{ris} phase shift selection problems from the vertices of $M$-dimensional probability simplex to the simplex itself.
	The original problem is highly non-convex and we propose a \gls{bcd} algorithm that iteratively updates $\{\boldsymbol{p}_k\}_{k \in \mathcal{K}}$, $\boldsymbol{w}$ and $\boldsymbol{t}$.
